---
title: ""
format:
  html:
    page-layout: full
    css: styles.css
editor: visual
fig-width: 12  
---

::::: {style="display: flex; justify-content: space-between; align-items: center; margin-bottom: 2em; margin-top: 1em;"}
::: {style="flex-grow: 1;"}
# Congenital Aortic Stenosis
:::

::: {style="display: flex; gap: 1em; align-items: center;"}
![](www/hds_logo.svg){width="80px"} ![](www/ohdsi_logo.svg){width="80px"}
:::
:::::

```{r setup_libraries, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(omopgenerics)
library(CohortSurvival)
library(CohortCharacteristics)

study_data_path <- c("data/studyData.RData", "../data/studyData.RData")
study_data_path <- study_data_path[file.exists(study_data_path)][1]

if (is.na(study_data_path)) {
  stop("Could not find `studyData.RData` in `data/` or `../data/`.")
}

load(study_data_path)

if (!all(c("summarise_characteristics", "survival") %in% names(data))) {
  stop("Could not find `summarise_characteristics` and `survival` in `studyData.RData`.")
}
```

```{r prepare_results, echo=FALSE, warning=FALSE, message=FALSE}
label_cohort <- function(x) tools::toTitleCase(gsub("_congenital$", "", gsub("_", " ", x)))
label_outcome <- function(x) tools::toTitleCase(gsub("_", " ", sub("_cohort$", "", x)))

baseline_result <- data[["summarise_characteristics"]]
survival_result <- data[["survival"]]

index_cohort <- baseline_result |>
  filter(
    variable_name == "Number subjects",
    estimate_name == "count",
    strata_name == "overall"
  ) |>
  mutate(n = suppressWarnings(as.numeric(estimate_value))) |>
  filter(!is.na(n)) |>
  slice_max(order_by = n, n = 1, with_ties = FALSE) |>
  pull(group_level)

if (length(index_cohort) == 0 || is.na(index_cohort)) {
  stop("Could not infer the baseline index cohort from `studyData.RData`.")
}

baseline_characteristics_result <- baseline_result |>
  omopgenerics::filterGroup(group_level == index_cohort)

preindex_conditions <- baseline_characteristics_result |>
  filter(
    strata_name == "overall",
    grepl("Comorbidity cohorts", variable_name),
    estimate_name == "percentage",
    estimate_value != "-"
  ) |>
  transmute(
    condition = variable_level,
    percentage = as.numeric(estimate_value)
  )

top_preindex_condition_levels <- preindex_conditions |>
  slice_max(order_by = percentage, n = 6, with_ties = FALSE) |>
  pull(condition)

top_preindex_condition_order <- preindex_conditions |>
  slice_max(order_by = percentage, n = 6, with_ties = FALSE) |>
  arrange(percentage) |>
  pull(condition)

all_preindex_condition_order <- preindex_conditions |>
  arrange(percentage) |>
  pull(condition)

compact_baseline_summary_gt <- baseline_characteristics_result |>
  filter(
    strata_name == "overall",
    variable_name %in% c("Number subjects", "Age", "Age group", "Sex") |
      (variable_name == "Comorbidity cohorts flag -inf to 0" & variable_level %in% top_preindex_condition_levels)
  ) |>
  mutate(group_level = label_cohort(group_level)) |>
  CohortCharacteristics::tableCharacteristics(type = "gt", header = c("cdm_name", "group_level"))

age_group_levels <- baseline_characteristics_result |>
  filter(
    strata_name == "overall",
    variable_name == "Age group",
    estimate_name == "count"
  ) |>
  pull(variable_level)

age_group_levels <- age_group_levels[order(as.numeric(sub(" .*", "", age_group_levels)))]

baseline_age_sex_data <- baseline_characteristics_result |>
  filter(
    strata_name == "age_group",
    variable_name == "Sex",
    estimate_name == "count"
  ) |>
  transmute(
    age_group = factor(strata_level, levels = age_group_levels),
    sex = variable_level,
    count = as.numeric(estimate_value)
  )

top_preindex_conditions <- preindex_conditions |>
  slice_max(order_by = percentage, n = 10, with_ties = FALSE) |>
  arrange(percentage)

all_preindex_characteristics_by_age <- baseline_characteristics_result |>
  filter(
    strata_name == "age_group",
    grepl("Comorbidity cohorts", variable_name),
    estimate_name == "percentage"
  ) |>
  transmute(
    condition = factor(variable_level, levels = all_preindex_condition_order),
    age_group = factor(strata_level, levels = age_group_levels),
    percentage = dplyr::if_else(estimate_value == "-", 0, as.numeric(estimate_value))
  )

preindex_condition_ranges <- all_preindex_characteristics_by_age |>
  group_by(condition) |>
  summarise(
    min_percentage = min(percentage, na.rm = TRUE),
    max_percentage = max(percentage, na.rm = TRUE),
    .groups = "drop"
  )

survival_overall_result <- survival_result |>
  omopgenerics::filterStrata(strata_name == "overall") |>
  mutate(
    group_level = label_cohort(group_level),
    variable_level = label_outcome(variable_level)
  )

survival_age_result <- survival_result |>
  omopgenerics::filterStrata(strata_name == "age_group") |>
  mutate(
    group_level = label_cohort(group_level),
    variable_level = label_outcome(variable_level),
    strata_level = factor(strata_level, levels = age_group_levels)
  )

survival_age_sex_result <- survival_result |>
  omopgenerics::filterStrata(strata_name == "age_group &&& sex") |>
  mutate(
    group_level = label_cohort(group_level),
    variable_level = label_outcome(variable_level)
  )

```

## Baseline Summary

### Baseline Summary Table

```{r baseline_summary_table, echo=FALSE, warning=FALSE, message=FALSE}
compact_baseline_summary_gt
```

### Age Group by Sex

```{r baseline_age_by_sex, echo=FALSE, warning=FALSE, message=FALSE, fig.width=10, fig.height=6, fig.cap="Distribution of children with congenital aortic stenosis across age groups and sex in the index cohort at Great Ormond Street Hospital."}
ggplot(baseline_age_sex_data, aes(x = age_group, y = count, fill = sex)) +
  geom_col(position = "stack", width = 0.75) +
  labs(
    x = "Age group",
    y = "Children",
    title = "Age and Sex Distribution at Diagnosis",
    fill = "Sex"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 10),
    legend.position = "right"
  )
```

### Top Pre-Index Conditions

```{r top_preindex_conditions, echo=FALSE, warning=FALSE, message=FALSE, fig.width=10, fig.height=7, fig.cap="Most common pre-index comorbid conditions in the congenital aortic stenosis cohort, ranked by the percentage of children with each condition."}
ggplot(top_preindex_conditions, aes(x = reorder(condition, percentage), y = percentage)) +
  geom_col(fill = "#2C7FB8", width = 0.75) +
  coord_flip() +
  labs(
    x = NULL,
    y = "Children (%)",
    title = "Most Common Pre-Index Conditions"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 10)
  )
```

### Baseline Comorbidities by Age Group

```{r baseline_comorbidities_by_age, echo=FALSE, warning=FALSE, message=FALSE, fig.width=13, fig.height=7.5, fig.cap="Pre-index comorbidities by age group in the congenital aortic stenosis cohort, with each line showing the range across age groups for a given condition. Suppressed estimates based on counts under 5 are shown as 0%."}
ggplot(all_preindex_characteristics_by_age, aes(x = percentage, y = condition, color = age_group)) +
  geom_segment(
    data = preindex_condition_ranges,
    aes(x = min_percentage, xend = max_percentage, y = condition, yend = condition),
    inherit.aes = FALSE,
    linewidth = 0.55,
    color = "#D8E0E8"
  ) +
  geom_point(size = 3) +
  labs(
    x = "Percentage with diagnosis prior to or on index date (%)",
    y = NULL,
    title = "Baseline Comorbidities by Age Group",
    color = NULL,
    caption = "Suppressed estimates based on counts under 5 are displayed as 0% in this figure."
  ) +
  scale_color_manual(values = c("0 to 1" = "#1F5A91", "2 to 12" = "#2A9D8F", "13 to 17" = "#E07A5F")) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_line(color = "#E6EAF0", linewidth = 0.35),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 10.5),
    axis.text.y = element_text(size = 9),
    plot.title = element_text(size = 13, face = "bold"),
    plot.caption = element_text(size = 9, hjust = 0, color = "#4D4D4D"),
    plot.caption.position = "plot",
    plot.margin = margin(t = 8, r = 14, b = 8, l = 10),
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(size = 10)
  )
```

## Outcome Risk

### Overall Cumulative Failure by Outcome

```{r overall_cumulative_failure, echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=8, fig.cap="Kaplan-Meier cumulative failure curves for the main study outcomes in the congenital aortic stenosis cohort, shown overall for each outcome across follow-up."}
survival_overall_result |>
  CohortSurvival::plotSurvival(facet = c("outcome", "cdm_name"), cumulativeFailure = TRUE) +
  labs(
    x = "Time in days",
    y = "Cumulative failure",
    title = "Overall Cumulative Failure by Outcome"
  ) +
  facet_grid(
    outcome ~ cdm_name,
    scales = "free_x",
    switch = "y",
    labeller = labeller(outcome = as_labeller(label_outcome))
  ) +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "#E0E0E0", color = NA),
        strip.text.x = element_text(size = 12, face = "bold"),
        strip.text.y.left = element_text(size = 11, face = "bold", angle = 0),
        strip.placement = "outside",
        legend.position = "right", panel.grid.minor = element_blank(),
        axis.title = element_text(size = 11), axis.text = element_text(size = 10))
```

### Age-Stratified Cumulative Failure by Outcome

```{r age_stratified_cumulative_failure, echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=8, fig.cap="Kaplan-Meier cumulative failure curves for each study outcome stratified by age group in the congenital aortic stenosis cohort."}
if (nrow(survival_age_result) > 0) {
  survival_age_result |>
    CohortSurvival::plotSurvival(facet = c("outcome", "cdm_name"), colour = "age_group", cumulativeFailure = TRUE) +
    labs(
      x = "Time in days",
      y = "Cumulative failure",
      title = "Age-Stratified Cumulative Failure by Outcome",
      color = "Age Group",
      fill = "Age Group"
    ) +
    facet_grid(
      outcome ~ cdm_name,
      scales = "free_x",
      switch = "y",
      labeller = labeller(outcome = as_labeller(label_outcome))
    ) +
    theme_minimal() +
    theme(
      strip.background = element_rect(fill = "#E0E0E0", color = NA),
      strip.text.x = element_text(size = 11, face = "bold"),
      strip.text.y.left = element_text(size = 11, face = "bold", angle = 0),
      strip.placement = "outside",
      panel.grid.minor = element_blank(),
      axis.title = element_text(size = 11),
      axis.text = element_text(size = 10),
      legend.position = "right"
    )
} else cat("No age-stratified cumulative failure data available.")
```

## Supplementary Figures

### Age-Sex Cumulative Failure

```{r age_sex_cumulative_failure, echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=8, fig.cap="Kaplan-Meier cumulative failure curves for each study outcome stratified jointly by age group and sex. This figure is presented as supplementary because the panel layout is more detailed."}
if(nrow(survival_age_sex_result) > 0) {
  survival_age_sex_result |>
    CohortSurvival::plotSurvival(facet = c("outcome", "sex"), colour = "age_group", cumulativeFailure = TRUE) +
    labs(
      x = "Time in days",
      y = "Cumulative failure",
      title = "Age- and Sex-Stratified Cumulative Failure",
      color = "Age Group",
      fill = "Age Group"
    ) +
    facet_grid(
      outcome ~ sex,
      scales = "free_x",
      switch = "y",
      labeller = labeller(outcome = as_labeller(label_outcome))
    ) +
    theme_minimal() +
    theme(
      strip.background = element_rect(fill = "#E0E0E0", color = NA),
      strip.text.x = element_text(size = 11, face = "bold"),
      strip.text.y.left = element_text(size = 11, face = "bold", angle = 0),
      strip.placement = "outside",
      panel.grid.minor = element_blank(),
      axis.title = element_text(size = 11),
      axis.text = element_text(size = 10),
      legend.position = "right"
    )
} else cat("No age-sex cumulative failure data available.")
```

## Cohort Count

### Cohort distribution

```{r cohort_distribution, echo=FALSE, warning=FALSE, message=FALSE, fig.width=10, fig.height=6, fig.cap="Procedure combinations among children with congenital aortic stenosis across study cohorts."}
library(tidyr)
library(stringr)
library(patchwork)

procedure_columns <- c(
  "aortic_valve_repair",
  "surgical_valvotomy",
  "percutaneous_valvotomy",
  "ross_procedure",
  "aortic_valve_replacement"
)

res <- data[["summarise_cohort_count"]] |>
  filter(
    strata_level != "overall",
    variable_name == "Number subjects"
  ) |>
  tidy()

missing_procedure_columns <- setdiff(procedure_columns, names(res))

if (length(missing_procedure_columns) > 0) {
  stop(
    paste0(
      "The cohort count result is missing required procedure columns: ",
      paste(missing_procedure_columns, collapse = ", ")
    )
  )
}

res <- res |>
  mutate(across(all_of(procedure_columns), as.integer))
 
 
upsetProc <- function(data, working_cohort){

plot_data <- data |>

  filter(cohort_name == working_cohort)

if (nrow(plot_data) == 0) {
  return(ggplot() + theme_void() + labs(title = paste("No cohort count data for", working_cohort)))
}

proc_cols <- procedure_columns
 
df_sorted <- plot_data |>

  # filter(if_any(all_of(proc_cols), ~ .x == 1)) |>

  mutate(

    plot_count = if_else(is.na(count), 1, count),

    count_label = if_else(is.na(count), "<5", as.character(count))

  ) |>

  arrange(desc(plot_count)) |>

  mutate(combo_id = factor(row_number()))
 
p_top <- df_sorted |>

  ggplot(aes(x = combo_id, y = plot_count)) +

  geom_col(fill = "#2C3E50", width = 0.6) +

  geom_text(aes(label = count_label), vjust = -0.5, size = 3.5) +

  theme_minimal() +

  labs(y = "N subjects") +

  theme(

    axis.text.x = element_blank(),

    axis.title.x = element_blank(),

    axis.ticks.x = element_blank(),

    panel.grid.major.x = element_blank(),

    panel.grid.minor.x = element_blank(),

    plot.margin = margin(b = 0) 

  )
 
df_matrix <- df_sorted |>

  select(combo_id, all_of(proc_cols)) |>

  pivot_longer(

    cols = all_of(proc_cols),

    names_to = "procedure",

    values_to = "is_active"

  ) |>

  mutate(

    procedure = str_replace_all(procedure, "_", " ") |> str_to_title(),

    procedure = factor(procedure, levels = rev(str_replace_all(proc_cols, "_", " ") |> str_to_title()))

  )
 
p_bottom <- df_matrix |>

  ggplot(aes(x = combo_id, y = procedure)) +

  geom_line(

    data = df_matrix , 

    aes(group = combo_id), 

    color = "black", 

    linewidth = 1

  ) +

  geom_point(aes(color = factor(is_active)), size = 4) +

  scale_color_manual(values = c("0" = "grey85", "1" = "black")) +

  theme_minimal() +

  theme(

    legend.position = "none",

    axis.text.x = element_blank(),

    axis.title.x = element_blank(),

    axis.title.y = element_blank(),

    panel.grid.major.x = element_blank(),

    panel.grid.minor = element_blank(),

    plot.margin = margin(t = 0) 

  )
 
upset_plot <- p_top / p_bottom + 

  plot_layout(heights = c(3, 1.5))+

  plot_annotation(

    title = paste0("Procedure Combinations for ", working_cohort),

    theme = theme(

      plot.title = element_text(hjust = 0.5, face = "bold", size = 14)

    )

  )
 
upset_plot
 
}

upsetProc(res, "aortic_stenosis_congenital")

upsetProc(res, "as_no_complex_chd")

upsetProc(res, "isolated_as")

upsetProc(res, "multi_level_as")

 
 
```
